---
title: "Chap 2: Linear Regression"
author: "Seong yeon Park"
date: "2025-01-24"
institute: Department of Statistic \newline Sungshin Women's University
fonttheme: "serif"
fontsize: 8pt
output:
  beamer_presentation:
    latex_engine: xelatex
    theme: "metropolis"
header-includes:
  - \input{header_includes.tex}
---

# Least Squares Method

## Simple Linear Regression

\begin{itemize}
    \item The data consists of $(x_1,y_1),...,(x_N,y_N)$
    $$
    {y_i} = \beta_0+\beta_1{x_i}+ \varepsilon_i
    $$
    \begin{itemize}
        \item $\beta_0$: intercept
        \item $\beta_1$: slope
        \item $\varepsilon_i$: random error
    \end{itemize}
\end{itemize}

- We obtain $\beta_0$ and $\beta_1$ via the least squares method.


## Least Squares Method

- sum of squares of the residuals,\newline we minimize $L$ of the squared distances $L$ between $(x_i,y_i)$ and $(x_i,\beta_0 + \beta_1{x_i})$ over $i = 1,2, ...,N$.
\begin{align*}
L=\sum_{i=1}^N(y_i-\beta_0-\beta_1{x_i})^2
\end{align*}

- Then, by partially differentiating $L$ by $\beta_0,\beta_1$ and letting them be zero.
\begin{align*}
\frac{\partial L}{ \partial\beta_0} &=-2\sum_{i=1}^N(y_i-\beta_0-\beta_1{x_i})=0 \\
\frac{\partial L}{ \partial\beta_1} &= -2\sum_{i=1}^N(x_i(y_i-\beta_0-\beta_1{x_i}))=0
\end{align*}
- $\beta_0$ and $\beta_1$ are regarded as constants when differentiating $L$ by $\beta_1$ and $\beta_0$.

## Least Squares Method

- When $\sum_{i=1}^N(x_i- \bar x)^2$ $\ne$ 0,
\newline $\hat\beta_0$, $\hat\beta_1$ instead of $\beta_0$, $\beta_1$ which means that they are not the true values but rather estimates obtained from data.
\begin{align*}
\hat\beta_1 &=\frac{\sum_{i=1}^N(x_i- \bar x)(y_i- \bar y)}{\sum_{i=1}^N(x_i- \bar x)^2} \\
\hat\beta_0 &=\bar y - \hat\beta_1{\bar x}
\end{align*}

- We center the data as follows,
\begin{align*}
\tilde{x}_1 := x_1 - \bar x, \cdots,\tilde{x}_N := x_N- \bar x, \tilde{y}_1 := y_1 - \bar y, \cdots,\tilde{y}_N := y_N- \bar y
\end{align*}

- Center the data results in a zero average.

- The formula for calculating the slope from the centralized data is as follows:
\begin{align*}
\hat\beta_1 &=\frac{\sum_{i=1}^N \tilde{x}_i\tilde{y}_i}{\sum_{i=1}^N(\tilde{x}_i)^2}
\end{align*}

## Example

- The two lines $l$ is obtained from the $N$ pairs of data and the least squares method, and  $l'$ obtained by shifting $l$ so that it goes through the origin.

```{r,echo=TRUE,eval=FALSE}
min.sq=function(x,y){
  x.bar=mean(x);y.bar=mean(y)
  beta.1=sum((x-x.bar)*(y-y.bar))/sum((x-x.bar)^2);beta.0=y.bar-beta.1*x.bar
  return(list(a=beta.0,b=beta.1))
}
a=rnorm(1);b=rnorm(1);
N=100;x=rnorm(N);y=a*x+b+rnorm(N)
plot(x,y);abline(h=0);abline(v=0)
abline(min.sq(x,y)$a,min.sq(x,y)$b,col="red")
x=x-mean(x);y=y-mean(y)
abline(min.sq(x,y)$a,min.sq(x,y)$b,col="blue")
legend("topleft",c("BEFORE","AFTER"),lty=1,col=c("red","blue"))
```

## Example
```{r,echo=FALSE}
set.seed(60)
min.sq=function(x,y){
  x.bar=mean(x);y.bar=mean(y)
  beta.1=sum((x-x.bar)*(y-y.bar))/sum((x-x.bar)^2);beta.0=y.bar-beta.1*x.bar
  return(list(a=beta.0,b=beta.1))
}
a=rnorm(1);b=rnorm(1);
N=100;x=rnorm(N);y=a*x+b+rnorm(N)
plot(x,y);abline(h=0);abline(v=0)
abline(min.sq(x,y)$a,min.sq(x,y)$b,col="red")
x=x-mean(x);y=y-mean(y)
abline(min.sq(x,y)$a,min.sq(x,y)$b,col="blue")
legend("topleft",c("BEFORE","AFTER"),lty=1,col=c("red","blue"))
```

# Multiple Regression

## Multiple Regression with Matrices

We formulate the least squares method for multiple regression with matrices. 

- $L:=\sum_{i=1}^N(y_i-\beta_0-\beta_1{x_i})^2$,
\begin{align*}
L = \parallel y-X\beta \parallel^2 = (y-X\beta)^T(y-X\beta)
\end{align*}

- If we define,
\begin{align*}
y :=
\begin{bmatrix}
y_1 \\
\vdots \\
y_N
\end{bmatrix}
, X :=
\begin{bmatrix}
1 & x_{1,1} &\cdots &x_{1,p}\\
\vdots & \vdots &\ddots &\vdots \\
1 & x_{N,1} &\cdots &x_{N,p}
\end{bmatrix}
, \beta :=
\begin{bmatrix}
\beta_0 \\
\beta_1 \\
\vdots \\
\beta_p
\end{bmatrix}
\end{align*}

- Partial differentiation with $L$
\begin{align*}
\nabla L := 
\begin{bmatrix}
\frac{\partial L}{\partial\beta_0} \\
\frac{\partial L}{\partial\beta_1}
\end{bmatrix}
= -2X^T(y-X\beta)
\end{align*}

## Multiple Regression

- Set to zero to find the minimum value
\begin{align*}
-2X^T(y-X\beta) =
\begin{bmatrix}
-2\sum_{i=1}^N(y_i-\sum_{j=0}^p \beta_j x_{i,j}) \\
-2\sum_{i=1}^N x_{i,1}(y_i-\sum_{j=0}^p \beta_j x_{i,j}) \\
\vdots \\
-2\sum_{i=1}^N x_{i,p}(y_i-\sum_{j=0}^p \beta_j x_{i,j})
\end{bmatrix}
\end{align*}

- When a matrix $X^TX$ is invertible, we have
\begin{align*}
\hat\beta= (X^TX)^{-1} X^Ty
\end{align*}


## When $X^TX$ is irreversible
1. $N < p+1$
\begin{align*}
rank(X^TX) \le rank(X) \le min \{N, p+1\} = N < p+1
\end{align*}
If $N > p$, It is $X$_particular, So there is no inverse matrix. 

2. Two columns in $X$ coincide.
\begin{align*}
X^TX z=0 \Rightarrow z^TX^TX_Z=0 \Rightarrow \parallel X_z \parallel^2 =0 \Rightarrow X_z=0
\end{align*}

# Distribution of $\hat\beta$

## Distribution of $\hat\beta$ 제목 생각해보기
- $y$ have been obtained from the covariates $X$ multiplied by the (true) coefficients $\beta$ plus some noise $\epsilon$.
\begin{align*}
y = X\beta +  \epsilon
\end{align*}

- The true $\beta$ is unknown and different from the estimate $\hat\beta$.
- We have estimated $\hat\beta$ via the least squares method from the $N$ pairs of data ($x_1,y_1$),$\cdots$,($x_N,y_N$) $\in$ $R^p$ X $R$

- $x_i$ $\in$ $R^p$ is the row vector consisting of $p$ values excluding the leftmost one in the $i$th row of $X$. 


## Density function
- We assume that each element $\epsilon_1,\cdots,\epsilon_N$ in the random variable $\epsilon$ is independent of the others and Gaussian distribution with mean zero and variance $\sigma^2$. $N(0,\sigma^2)$
\begin{align*}
f_i(\epsilon_i) = \frac{1}{\sqrt {2\pi\sigma^2}} e^{-\frac{\epsilon_i^2}{2\sigma^2}}
\end{align*}

- We may express the distributions of $\epsilon_1,\cdots,\epsilon_N$ by
\begin{align*}
f(\epsilon) = \prod_{i=1}^N f_i(\epsilon_i) = \frac{1}{(2\pi\sigma^2)^{N/2}} e^{-\frac{\epsilon^T \epsilon}{2\sigma^2}}
\end{align*}
This is $N(0,\sigma^2I)$, $I$ is a unit matrix of size $N$.


## Independent if and only if their covariance is zero

- For the proof,
\begin{align*}
\hat\beta = (X^TX)^{-1}X^T(X\beta+\epsilon) = \beta + (X^TX)^{-1}X^T\epsilon
\end{align*}

- Since the average of $\epsilon \in R^N$ is zero, the average of $\epsilon$ multiplied from left by the constant matrix $(X^TX)^{-1}X^T$ is zero.
\begin{align*}
E[\hat\beta] = \beta
\end{align*}

- In general, we say that an estimate is unbiased if its average coincides with the true value.

## Covariance matrix of $\hat\beta$

- $\hat\beta$ and its average $\beta$ consist of $p+1$ values. 
- $V(\hat\beta_i) = E(\hat\beta_i - \beta_i)^2, i=0,1,\cdots,p$, the covariance $\sigma_{i,j} := E(\hat\beta_i - \beta_i)(\hat\beta_j - \beta_j)^T$ can be defined for each pair $i\ne j$.

- matrix consisting of $\sigma_{i,j}$ in the $i$th row and $j$th column as to the covariance matrix of $\hat\beta$.

\begin{align*}
& E
\begin{bmatrix}
(\hat\beta_0 - \beta_0)^2 & (\hat\beta_0 - \beta_0)(\hat\beta_1 - \beta_1) &\cdots &(\hat\beta_0 - \beta_0)(\hat\beta_p - \beta_p)\\
(\hat\beta_1 - \beta_1)(\hat\beta_0 - \beta_0) &(\hat\beta_1 - \beta_1)^2 &\cdots &(\hat\beta_1 - \beta_1)(\hat\beta_p - \beta_p)\\
\vdots &\vdots &\ddots &\vdots \\
(\hat\beta_p - \beta_p)(\hat\beta_0 - \beta_0) &(\hat\beta_p - \beta_p)(\hat\beta_1 - \beta_1) &\cdots &(\hat\beta_p - \beta_p)^2
\end{bmatrix} 
\end{align*}

## Covariance matrix of $\hat\beta$

\begin{align*}
& E
\begin{bmatrix}
(\hat\beta_0 - \beta_0)^2 & (\hat\beta_0 - \beta_0)(\hat\beta_1 - \beta_1) &\cdots &(\hat\beta_0 - \beta_0)(\hat\beta_p - \beta_p)\\
(\hat\beta_1 - \beta_1)(\hat\beta_0 - \beta_0) &(\hat\beta_1 - \beta_1)^2 &\cdots &(\hat\beta_1 - \beta_1)(\hat\beta_p - \beta_p)\\
\vdots &\vdots &\ddots &\vdots \\
(\hat\beta_p - \beta_p)(\hat\beta_0 - \beta_0) &(\hat\beta_p - \beta_p)(\hat\beta_1 - \beta_1) &\cdots &(\hat\beta_p - \beta_p)^2
\end{bmatrix} \\
&= E
\begin{bmatrix}
\hat\beta_0 - \beta_0 \\
\hat\beta_1 - \beta_1 \\
\vdots \\
\hat\beta_p - \beta_p
\end{bmatrix}
\begin{bmatrix}
\hat\beta_0 - \beta_0, \hat\beta_1 - \beta_1, \cdots, \hat\beta_p - \beta_p
\end{bmatrix} \\
&= E(\hat\beta-\beta)(\hat\beta-\beta)^T = E(X^TX)^{-1}X^T \epsilon{(X^TX)^{-1}X^T\epsilon}^T \\
&= (X^TX)^{-1}X^T E \epsilon\epsilon^T X(X^TX)^{-1} = \sigma^2(X^TX)^{-1}
\end{align*}

We have determined that the covariance matrix of $\epsilon$ is $E\epsilon\epsilon^T =\sigma^2I$.

\begin{align*}
\hat\beta \sim N(\beta,\sigma^2(X^TX)^{-1})
\end{align*}

