---
title: 'Chapter 6 : Regularization'
author: "Seong yeon Park"
date: "March 19, 2025"
output:
  beamer_presentation:
    latex_engine: xelatex
    theme: metropolis
  ioslides_presentation: default
institute: The Three Sisters of Newton \newline School of Mathematics, Statistics
  and Data Science \newline Sungshin Women’s University
fonttheme: serif
fontsize: 8pt
subtitle: 6.1 Ridge
header-includes: \input{header_includes.tex}
---

# Ridge

## Chapter 2: Linear Regression

We formulate the least squares method for multiple regression with matrices. 
\begin{align*}
L:&=\sum_{i=1}^N(y_i-\beta_0-\beta_1{x_{i,1}},\cdots,-\beta_1{x_{i,p}})^2,\\
L &= \parallel y-X\beta \parallel^2 = (y-X\beta)^T(y-X\beta)\\
&=y^Ty-y^TX\beta-\beta^TX^Ty-\beta^TX^TX\beta
\end{align*}

- Partial differentiation with $L$
\begin{align*}
\nabla 
L:=
\frac{\partial L}{ \partial\beta}
= -X^Ty-X^Ty+2X^TX\beta = -2X^T(y-X\beta)
\end{align*}

- Set to zero to find the minimum value
\begin{align*}
-2X^T(y-X\beta) = 0
\end{align*}


## Multiple Regression Singular/Nonsingular Matrix
- When a matrix $X^TX$ is nonsingular, we have
\begin{align*}
2X^TX\beta=2X^Ty \\
\hat\beta= (X^TX)^{-1} X^Ty
\end{align*}

- If matrix $X^TX$ is singular, Determinant is too small, $\beta$ becomes large and an inconvenient situation occurs.

## Ridge Regression

- $\lambda$ $\ge$ 0 be a constant, we often use to minimize the square error plus by the squared norm of $\beta$ multiplied by $\lambda$.

- Loss function of existing linear regression (sum of error squared)
\begin{align*}
L = \frac{1}{N} \parallel y-X\beta \parallel^2
\end{align*}

- Loss function of ridge regression
\begin{align*}
L := \frac{1}{N} \parallel y-X\beta \parallel^2 + \lambda \parallel \beta \parallel_2^2
\end{align*}

- $\lambda\parallel \beta \parallel^2_2$ is the regularization term of ridge regression.
- The larger the $\lambda$, the smaller the $\beta$ size.


is the square of the L2 norm of $\beta$.

## Differentiate Loss function of Ridge Regression
- Loss function of ridge regression
\begin{align*}
L &:= \frac{1}{N} \parallel y-X\beta \parallel^2 + \lambda \parallel \beta \parallel_2^2 \\
&= \frac{1}{N}(y-X\beta)^T(y-X\beta) + \lambda\beta^T\beta \\
&= \frac{1}{N}(y^Ty-y^TX\beta - \beta^TX^Ty+\beta^TX^TX\beta)+ \lambda\beta^T\beta
\end{align*}

- Differentiate L by $\beta$,
\begin{align*}
\frac{\partial L}{ \partial\beta} &= \frac{1}{N}(-2\beta^TX^Ty+2X^TX\beta+2\lambda\beta) \\
&= -\frac{2}{N}X^T(y-X\beta) +2\lambda\beta = 0 \\
&=-\frac{1}{N}X^T(y-X\beta)+\lambda\beta = 0 \\
&=\frac{1}{N}X^T(y-X\beta)=\lambda\beta
\end{align*}

- This additional term serves to control the size of $\beta$.

## Calculate the Weight $\hat\beta$

- If $X^TX+\lambda I$ is nonsingular,
\begin{align*}
\frac{1}{N}X^T(y-X\beta)=\lambda\beta \\
X^T(y-X\beta)=N\lambda\beta \\
X^Ty-X^TX\beta = N\lambda\beta \\
X^Ty=X^TX\beta+N\lambda\beta \\
X^Ty=(X^TX+N\lambda I)\beta
\end{align*}
- $\hat\beta=(X^TX+N\lambda I)^{-1}X^Ty$



## R Code for Ridge Regression
```{r}
ridge=function(X,y,lambda=0){
  X=as.matrix(X);p=ncol(X);n=length(y);X.bar=array(dim=p);s=array(dim=p)
  for (j in 1:p){X.bar[j]=mean(X[,j]);X[,j]=X[,j]-X.bar[j];};
  for (j in 1:p){s[j]=sd(X[,j]);X[,j]=X[,j]/s[j]};
  y.bar=mean(y);y=y-y.bar
  beta=drop(solve(t(x)%*%X+n*lambda*diag(p))%*%t(X)%*%y)
  for (j in 1:p)beta[j]=beta[j]/s[j]
  beta.0=y.bar-sum(X.bar*beta)
  return(list(beta=beta,beta.0=beta.0))
}
```

## Example 48
```{r,eval=FALSE}
df=read.table("crime.txt");x=df[,3:7];y=df[,1];p=ncol(x);
lambda.seq=seq(0,100,0.1);coef.seq=lambda.seq
plot(lambda.seq,coef.seq,xlim=c(0,100),ylim=c(-40,40),
     xlab="lambda",ylab="beta",main="The coefficients for each lambda",
     type="n",col="red")
for (j in 1:p){
  coef.seq=NULL;for(lambda in lambda.seq)coef.seq=c(coef.seq,
                                                    ridge(x,y,lambda)$beta[j])
  par(new=TRUE);lines(lambda.seq,coef.seq,col=j)
}
legend("topright",legend=
         c("annual police funding in $resident","% of people 25 years +
           with 4 yrs. of high school",
           "% of 16 to 19 year-olds not in highschool and not highschool 
           graduates", "% of 18 to 24 year-olds in college",
           "% of 18 to 24 year-olds in college"),col=1:p,lwd=2,cex=.8)
```

## Example 48

```{r,echo=FALSE}
df=read.table("crime.txt");x=df[,3:7];y=df[,1];p=ncol(x);
lambda.seq=seq(0,100,0.1);coef.seq=lambda.seq
plot(lambda.seq,coef.seq,xlim=c(0,100),ylim=c(-40,40),
     xlab="lambda",ylab="beta",main="The coefficients for each lambda",
     type="n",col="red")
for (j in 1:p){
  coef.seq=NULL;for(lambda in lambda.seq)coef.seq=c(coef.seq,
                                                    ridge(x,y,lambda)$beta[j]) 
  par(new=TRUE);lines(lambda.seq,coef.seq,col=j)
}
legend("topright",legend=
         c("annual police funding in $resident","% of people 25 years + with 4 yrs. of high school", 
           "% of 16 to 19 year-olds not in highschool and not highschool graduates", 
           "% of 18 to 24 year-olds in college",
           "% of 18 to 24 year-olds in college"),col=1:p,lwd=2,cex=.8)
```


## 

\begin{center}
  {\bf {\Huge Q \& A}}
\end{center}


## 

\begin{center}
  {\bf {\Huge Thank You}}
\end{center}

